---
title: "Reproducible Research: Peer Assessment 1"
output:
  html_document:
    keep_md: true
---

## Loading and preprocessing the data

```{r chunk-1}
library(tidyverse)

file.name <- "activity.csv"
if(!file.exists(file.name)) {
    unzip("activity.zip")
}
steps <- read_csv(file.name, col_types = "iDi")
```

## What is mean total number of steps taken per day?

```{r chunk-2}
steps.per.day <- steps %>%
    group_by(date) %>%
    summarise(steps = sum(steps, na.rm = TRUE))
steps.per.day %>%
    ggplot(aes(x = steps)) +
    geom_histogram(binwidth = 1000, boundary = 0)
summary(steps.per.day$steps)
```



## What is the average daily activity pattern?

```{r chunk-3}
steps.per.interval <- steps %>%
    group_by(interval) %>%
    summarise(steps = mean(steps, na.rm = TRUE))
steps.per.interval %>%
    ggplot(aes(x = interval, y = steps)) +
    geom_line()
max.interval <- (steps.per.interval %>% filter(steps == max(steps)))$interval
```

The 5-minute interval `r max.interval`, on average across all the days in the dataset, contains the maximum number of steps.

## Imputing missing values

```{r chunk-4}
intervals.without.steps <- nrow(steps %>% filter(is.na(steps)))
intervals.with.steps <- nrow(steps)
```

In `r intervals.without.steps` of `r intervals.with.steps` intervals (`r round(intervals.without.steps / intervals.with.steps * 100, digits = 2)` %) the number of steps is missing.

We decided to substitute the missing values by the mean of the corresponding 5-minute interval:

```{r chunk-5}
substitute.step.values <- rep(as.integer(round(steps.per.interval$steps)), 61)
steps.imputed <- steps %>%
    mutate(steps = if_else(is.na(steps), substitute.step.values, steps))
steps.per.day.imputed <- steps.imputed %>%
    group_by(date) %>%
    summarise(steps = sum(steps))
steps.per.day.imputed %>%
    ggplot(aes(x = steps)) +
    geom_histogram(binwidth = 1000, boundary = 0)
summary(steps.per.day.imputed$steps)
```

After imputing missing data there are less days with less than 1000 steps.
We see that the mean and the median are higher than before.

## Are there differences in activity patterns between weekdays and weekends?

```{r chunk-6}
determine.day.type <- function(date) {
    if_else(as.POSIXlt(date)$wday %in% c(0, 6), "weekend", "weekday")
}
steps.imputed <- steps.imputed %>%
    mutate(type = as.factor(determine.day.type(date)))
steps.by.interval.and.type.imputed <- steps.imputed %>%
    group_by(interval, type) %>%
    summarise(steps = mean(steps), .groups = "drop_last")
steps.by.interval.and.type.imputed %>%
    ggplot(aes(x = interval, y = steps)) +
    geom_line() +
    facet_grid(vars(type))
```

On weekdays there are more steps in the morning.
On weekends there are more steps in the afternoon.
